### 2013-09-19 Reports/Report1/data.tex

\section{Feedback Data}
After every translation, a pre-selected corrector revises the translation. After reviewing, the corrector assesses the translation and leave a feedback on it. This feedback consists of a comment and number which evaluate the quality of translation from one to five (five is the best). These feedbacks are also categorized in different aspects of translation like content, grammar, style, accuracy and etc. For each translation, the value related to each criterion (price, time, text matching and cooperation) can be fetched from database.

Since the website is working from 2011, there are about 2000 feedback records in database which related to different translators and language pairs.

Some sample records of database are shown in Appendix A.