### 2013-09-19 Reports/Report1/data.tex
revise report1

\section{Translation Feedback Data}
After every translation, a pre-selected corrector revises the translation and assesses the translation quality. Then the corrector leaves a feedback on it which consists of a comment and a number from one to five (five is the best) which evaluate the quality of translation. These feedbacks are also categorized in different aspects of translation like content, grammar, style, accuracy and etc. For each translation, criterion values (price, time, text matching and cooperation) offered by translator can be fetched from database.

Since the website is working from 2011, there are about 2000 feedbacks in database. These records contain the assessment of different translators and language pairs in different criterion categories.

Some sample records of database are shown in Appendix A.