### 2013-12-24 docs/reports/finalreport/report.tex
starting project final report

\input{config}
\input{title}

\begin{document}

\maketitle
\begin{abstract}
We did experiments on three datasets of different size and with different characteristics using Regression Algorithms. Four different algorithms where used for the experiment and for this four we did experiments varying their arguments data scaling and imputation techniques for missing data. In the report we first give an overview over the algorithms used, then we interpret the results of our experiments for each dataset and do a comparison of the different algorithms. In the end we try to compare the results of all datasets and draw conclusions.
\end{abstract}

\vfill
\tableofcontents
\newpage

\input{introduction}
\newpage
\input{problem}
\newpage
\input{architecture}
\newpage
\input{design}
\newpage
\input{development}
\newpage
\input{future}
\newpage

\bibliographystyle{plain}
\bibliography{report}

\end{document}