### 2013-12-25 docs/reports/finalreport/development.tex

\section{System Development}
In this section, some special points related to development of different components are described.

\subsection{Order}
Ordering page is a full client-side page following the idea of SPA (Single Page Application) using KnockoutJS and BreezeJS. Complex business of the company for applying query, offering translators and purchasing order urges using design models and standards.
\subsection{Machine Learning's Data Accumulating}
In order to tackle the problem of translator's ranking, a data accumulating system is designed. Similar to Order, based on query document it offers three translators and stores client's choice. The data gathered by the system is used in a Machine Learning program using Learning to Rank algorithms in order to discover the underlying pattern for proposing the translators. The details of the project is described in Seminar lecture related to the project. In contrast to Order system, it eliminates translator's picture and name. It prevents bias in final answer by client (estimator) and puts emphasis on presented criteria.
\subsection{Text Matching}
Using Lucene translator's documents are stored based on their language. In order to achieve the final text matching value, the similarity between each document and query document is calculated. The problem here is the aggregation algorithm between the results. Finally, based on the results that are reported in Seminar lecture, GP2\cite{gp2} algorithm is implemented.
\subsection{Translation Memory}
As it's mentioned before, the aim of Translation Memory system is helping the translators to have an access to their previous-translated texts and applying them to their new translations. HunAlign is used to align the sentences of one text with the sentences of its corresponding text in another language. Although the library tends to be very error-prone specially for complex texts, adding dictionary and using rather simple texts lead to more reasonable results. At the end, Lucene is used for indexing and searching the sentences.