### 2013-12-25 docs/reports/finalreport/introduction.tex

\section{Introduction}
As a hub for professional translation services, Nativy \cite{nativy}(founded 2011 in Vienna) offers its customers top-quality translations at reasonable terms.

The company works exclusively with qualified translators into their native language. All Nativy translations are reviewed in accordance with EN 15038, the European Quality Standard for Translation Services, and translators have to provide evidence of the required qualifications stipulated in the standard by means of their résumé.

By providing the platform, regardless to geographical location of translators and clients, they can communicate and proceed the translation process. It eases translating text to even not-so-popular languages and also reduces the final cost.

The features regarding to the internship has been developed since Summer 2013.

In the following, first we address the problem. Then in system architecture as well as design patterns and development techniques are explained. At the end related work to the project is briefly met.