### 2013-12-25 docs/reports/finalreport/report.tex

\input{config}
\input{title}

\begin{document}

\maketitle
\begin{abstract}
The reports explains the efforts done by design and develop a translator
retrieval system. The system can be defined in domain of on-line translation. It aims to find an appropriate translator for a client and manages the process till the last step. Besides, by providing a translation memory system translators can apply their previous translated texts in their new ones. In the project, C\texttt{\#}.NET and MS SQL SERVER as main platform and Lucene.NET as text processor are used. It is designed based on DDD (Data Driven Design) model. Windows Azure is also selected as deployment environment.
\end{abstract}

\vfill
\tableofcontents
\newpage

\input{introduction}
\newpage
\input{problem}
\newpage
\input{architecture}
\newpage
\input{design}
\newpage
\input{development}
\newpage

\bibliographystyle{plain}
\bibliography{report}

\end{document}