### 2013-12-28 docs/seminar/abstract.tex
writing abstract

Expertise Retrieval, whose task is to suggest people with relevant expertise regarding to a query, has received increasing interest in recent years. In this paper, we proposed an Expert Retrieval platform narrowed down to Translators as experts. Issues and obstacles during design and development as well as acquired solutions and results are reported. Ranking the translators using Learning to Rank as well as documents' aggregation functions are two main issues which are specifically studied in the paper.