### 2013-12-29 docs/reports/finalreport/development.tex

\section{Development}
In this section, some special issues related to development of different components are described.

\subsection{Order}
Ordering page is a full client-side page following the idea of SPA (Single Page Application) using KnockoutJS and BreezeJS. Complex business of the company in applying query, offering translators and purchasing order urges using design models and standards.
\subsection{Machine Learning's Data Accumulating}
In order to tackle the problem of translator's ranking, a data accumulating system is designed. Similar to Order, based on query document it offers three translators and stores client's choice. The data gathered by the system is used in a Machine Learning program using Learning to Rank algorithms. It aims to discover the underlying pattern for proposing the translators. The details of the project is described in seminar lecture related to the project. In contrast to Order system, the system eliminates translator's picture and name which prevents bias in final answer by estimator and puts emphasis on the presented criteria.
\subsection{Text Matching}
Using Lucene, translator's documents are stored regarding to their languages. In order to achieve the final text matching value, the similarity between each document and query document is calculated. The main problem here is the way to  aggregate result scores. As the problem is addressed in Seminar lecture, GP2\cite{gp2} algorithm is implemented.
\subsection{Translation Memory}
As it's mentioned before, the aim of Translation Memory system is helping the translators to exploit their previous-translated texts in their new translations. In order to develop the system, first HunAlign library is used for aligning the sentences of one text with the sentences of its corresponding text in another language. Although the algorithm is very error-prone specially for complex texts, adding dictionary and using rather simple and straightforward texts lead to more reasonable results. At the end, Lucene indexes the text and provides searching abilities.