### 2013-12-29 docs/reports/finalreport/introduction.tex

\section{Introduction}
As a hub for professional translation services, nativy \cite{nativy}(founded 2011 in Vienna) offers its customers top-quality translations at reasonable terms.

The company works exclusively with qualified translators into their native language. All nativy translations are reviewed in accordance with EN $15038$, the European Quality Standard for Translation Services, and translators have to provide evidence of the required qualifications stipulated in the standard by means of their résumé.

Using the platform, regardless to geographical location of translators and clients, they can communicate and proceed a translation process. It aims to ease translating text to even not-so-popular languages as well as reduction of final cost.

In the following, first I describe system requirements. Then system architecture and design are explained. Tools and Technologies applied in the platform are Section \ref{sec:tools}. At the end, some special issues during system development are reported in the last section.