### 2014-01-11 docs/reports/finalreport/introduction.tex

\section{Introduction}
As a hub for professional translation services, nativy \cite{nativy}(founded 2011 in Vienna) offers its customers top-quality translations at reasonable terms.

The company works exclusively with qualified translators into their native language. All nativy translations are reviewed in accordance with EN $15038$, the European Quality Standard for Translation Services, and translators have to provide evidence of the required qualifications stipulated in the standard by means of their résumé.

Using the platform, regardless of geographical location of translators and clients, they can communicate and proceed with translation. It aims to ease translating text to even not-so-popular languages as well as reduction of final cost.

The developed platform mainly consists of two parts. First, storing profiles of translators by providing registration as well as scheduling possibilities. Additionally, assisting them during translation process by offering a simple translation memory system. Second, providing a platform for clients to search and hire the translators and purchase the orders. Some complexities like text processing and expert ranking are tackled in the second part.

The remain of the report is organized as follows: In Section \ref{sec:scenario}, system requirements and usual scenarios are described in detail. Then in Section \ref{sec:architecture}, system architecture as well as design methods are explained. Applied tools and technologies following with a short description are introduced in Section \ref{sec:tools}. At the end in Section \ref{sec:development}, by demonstrating the final product in some screen-shots, specific issues in system development are reported.