### 2014-01-11 docs/reports/finalreport/report.tex

\newcommand{\figurescaling}{0.7}
\input{config}
\input{title}

\begin{document}

\maketitle
\begin{abstract}
The report explains the details of design and developing process of a translator retrieval platform. The platform can be defined in domain of on-line translation. It aims to suggest an appropriate translator to a client and manages the process of translation until the last step. Besides, by providing a translation memory system translators can exploit their previous translated texts in their new ones. C\texttt{\#}.NET and MS SQL SERVER as main platform and Lucene.NET as text processor are used. The system is designed based on DDD (Data Driven Design) model. Windows Azure is also selected as deployment environment.
\end{abstract}

\vfill
\tableofcontents
\newpage

\input{introduction}
\newpage
\input{problem}
\newpage
\input{architecture}
\newpage
\input{design}
\newpage
\input{development}
\newpage

\listoffigures
\listoftables
\bibliographystyle{plain}
\bibliography{report}

\end{document}