### 2014-01-11 docs/reports/finalreport/introduction.tex
adding new screen-shot

\section{Introduction}
As a hub for professional translation services, the platform offers its customers top-quality translations at reasonable terms. All translations are reviewed and have to provide evidence of the required qualifications stipulated in the standard by means of their résumé.

Using the platform, regardless of geographical location, translators and clients can communicate and proceed the translation. It aims to ease translating text to even not-so-popular languages as well as reduction of final cost.

The developed platform mainly consists of two parts. First, storing profiles of translators developed in registration and  scheduling systems. Additionally, assisting them during translation process by offering a simple translation memory system. Second, providing a platform for clients to search and hire the translators and purchase the orders.

The remain of the report is organized as follows: In Section \ref{sec:scenario}, system requirements and usual scenarios are described in detail. Then in Section \ref{sec:architecture}, system architecture as well as design methods are explained. Applied tools and technologies following with a short description are introduced in Section \ref{sec:tools}. At the end in Section \ref{sec:development}, by demonstrating the final product in some system screen-shots, specific issues in system development are reported.