### 2014-02-06 docs/reports/finalreport/architecture.tex
revising finalreport and paper

\section{Architecture and Design}
\label{sec:architecture}
The system is designed based on DDD (Data Driven Design) and Multi-Layer Architecture. As depicted in Figure \ref{fig:architecture}, the system consists of three main layers : Interface, Business and Data Access Layer. Data is passed between the layers using Model objects.

The interface layer is divided into two main parts. Server- and client-side. The server-side which is denoted as Application Layer is following the MVVC (Model-View, View, Controller) design model. Its main purpose is mapping Model-View objects of the Interface Layer to Model objects and passing them to the Business Layer and vice versa. Client-side web pages are designed based on the MVVM (Model-View, View, Model) model. Tools used for developing the model are explained in Section \ref{sec:tools}.

Using AOP (Aspect-Oriented Programming), logging, error handling and security systems are developed on top of other layers.

As well as the mentioned architectural design patterns, the following implementation design patterns are used:
\begin{itemize}
	\item Dependency Injection and Inversion of Control implemented in Interface Layer
	\item Singleton Repository in Data Access Layer
	\item Unit of Work for handling transactions
\end{itemize}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.8]{figures/architecture.jpg}
\caption{System Architecture
\label{fig:architecture}}
\end{center}
\end{figure}