### 2014-02-06 docs/reports/finalreport/introduction.tex

\section{Introduction}
As a hub for professional translation services, the platform offers its customers top-quality translations at reasonable terms. All translators have to provide evidence of the required qualifications by means of their résumé.

Using the platform, regardless of geographical location, translators and clients can communicate and produce the translation. The platform facilitates translating text to even not-so-popular languages.

The developed platform can mainly be divided into two parts. First, translator platform which consists of registration,  scheduling and translation memory system. Second, client platform which facilitates searching and hiring the translators as well as purchasing the orders.

The remainder of the report is organized as follows: In Section \ref{sec:scenario}, system requirements and usual scenarios are described in detail. Then in Section \ref{sec:architecture}, the system architecture as well as design methods are explained. Section \ref{sec:tools} introduces the tools and technologies we applied. Section \ref{sec:development} reports specific issues in the development of the system by demonstrating the final product in a series of system screen-shots.