### 2014-02-11 docs/seminar/evaluation.tex
adding significance test results

\section{Apply The Methods and Results}
\label{sec:apply}
In this section, we applied different approaches on the platform. By comparing the methods, we aim to discover the most appropriate one regarding to the project's characteristics and data.

\subsection{Aggregation Functions}
In order to compare different aggregation functions, similar to \citet{agg-gp2} three algorithms are selected. $GP2$ as well as $Top1$ and $Top5$ which are two common forms of $TopN$ aggregation algorithm. $TopN$ refers to algorithm that summarizes the $N$ top documents (i.e. $Top1$ only using the top associated document to rank the candidates).

Feedbacks of proof-readers after every translation are used as a basis for evaluating the algorithms. Since feedbacks are a measure for quality of translation, the more similar the ranking of algorithms to feedbacks are the better it is.

In order to calculate the correlation value, we applied Spearman Rank Order and Kendall Rank Correlation as two common methods. Table \ref{table:correlation} shows the results of Spearman correlation coefficient ($r_s$) and Kendall's tau coefficient ($\tau$) using $181$ records of purchased orders. Table \ref{table:signtest} represents the Significance Test of both methods calculated by R.

As it is usual in many cases, the coefficient value of Spearman is slightly higher than Kandell's. The outcome shows an approximately weak correlation between aggregation functions and feedbacks of proof-readers. TODO about significance test??.

Comparing the algorithms, $GP2$ outperforms the others in both correlation tests. In comparison to $Top1$, $Top5$ has slightly better performance. The results are also nearly the same when comparing based on language-pairs.

\begin{table}
\begin{center}
\scalebox{1.2}{
\begin{tabular}{p{1cm}|c|c|c|}
\cline{2-4} & Top1 & Top5 & GP2  \\
\hline \multicolumn{1}{ |c| }{$r_s$} & 0.052 & 0.089 & 0.145\\
\hline \multicolumn{1}{ |c| }{$\tau$} & 0.034 & 0.059 & 0.102\\
\hline
\end{tabular}
}
\caption{Correlation Test Between Algorithms and Proof-readers' Feedbacks}
\label{table:correlation}
\end{center}
\end{table}

\begin{table}
\begin{center}
\scalebox{1.2}{
\begin{tabular}{p{1cm}|c|c|c|}
\cline{2-4} & Top1 & Top5 & GP2  \\
\hline \multicolumn{1}{ |c| }{$r_s$} & 0.4866 & 0.2295 & 0.05038\\
\hline \multicolumn{1}{ |c| }{$\tau$} & 0.5157 & 0.2562 & 0.05263\\
\hline
\end{tabular}
}
\caption{P-Value of Significance of Correlation Test Algorithms}
\label{table:signtest}
\end{center}
\end{table}

\subsection{Learning To Rank}
TODO