### 2014-04-14 docs/seminar/abstract.tex
changed abstract and others

%Expertise Retrieval, whose task is to suggest people with relevant expertise regarding to a query, has received increasing interest in recent years. In this paper, we propose an Expert Retrieval platform narrowed down to translators as experts. Issues and obstacles during design and development as well as acquired solutions and results are reported. Ranking the translators using Learning to Rank as well as documents' aggregation functions are two main issues which are specifically studied in the paper.

This article describes a method and tool to identify expert translators in an online translation service. We start from existing efforts on expert retrieval and factor in additional parameters based on the real-world scenario of the task. The system first identifies topical expertise using an aggregation function over relevance scores of previously translated documents by each translator, and then a learning to rank method to factor in non-topical relevance factors that are part of the decision-making process of the user, such as price and duration of translation. We test the system on a manually created test collection and show that the method is able to effectively support the user in selecting the best translator.